---
title: "Chemical space covered by bitter tastants, food derived compounds and drug compounds"
subtitle: "Supplementary figures to Dunkel, Hofmann and Di Pizio (2020): In-silico investigation of bitter hop-derived compounds and their cognate bitter taste receptors"
  
output:
  html_document:
      toc: true
      theme: united
      toc_float: true
---



```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

```{r load packages,warning=FALSE, message=FALSE }
library(here)
library(tidyverse)
library(cowplot)
library(ggrepel)
library(RColorBrewer)
library(plotly)
library(kableExtra)
library(DT)
library(ggforce)
library(stringi)

theme_set(theme_cowplot())
```

```{r}

daten <- read.csv(here("data", "daten_plot4_neu.csv"), dec = ".", sep = ",", header = TRUE)
daten2 <- daten %>%
          distinct(isomeriSmiles, .keep_all = TRUE)

daten2$superclass2 <- ordered(daten2$superclass2, levels = c("Others", "Phenylpropanoids and polyketides", "Organosulfur compounds", "Organoheterocyclic compounds", 
                                                             "Organic oxygen compounds", "Organic nitrogen compounds", "Organic acids and derivatives", 
                                                             "Lipids and lipid-like molecules", "Benzenoids", "Alkaloids and derivatives")) 
```

# Chemical space of bitter tastants

### t-SNE plot of the bitter chemical space, coloured according FooDB and DrugBank co-occurrence

```{r, fig.width=12, fig.height=12}
p <- plot_ly(daten2, x = ~X1, y = ~X2, z = ~X3, color = ~group.y,
             colors = c("#ECE471",   "#ED7F85", "#C0D4A4",  "#BD899F")) %>%
  add_markers(marker = list(size = 5, opacity = 1), hoverinfo = "text",
              text = ~paste("Compound Name: ", str_wrap(IUPACName, width = 40),
                            '<br> PubChem CID: ', cid2,
                            '<br> Database: ', group.y
                            )) %>%
  layout(scene = list(xaxis = list(title = 'tSNE1'),
                      yaxis = list(title = 'tSNE2'),
                      zaxis = list(title = 'tSNE3')),
         legend = list(itemsizing = "constant"))
p
```






### t-SNE plot of the bitter chemical space, coloured according to ClassyFire chemical superclasses

```{r, fig.width=12, fig.height=12}
p <- plot_ly(daten2, x = ~X1, y = ~X2, z = ~X3, color = ~superclass2, 
             colors = c("grey79", "#b15928", "#6a3d9a", "#cab2d6", "#ff7f00", "#e31a1c", "#33a02c", "#1f78b4", "#b2df8a", "#a6cee3")) %>%
              add_markers(marker = list(size = 5, opacity = 1), hoverinfo = "text",
              text = ~paste("Compound Name: ", str_wrap(IUPACName, width = 40),
                            '<br> PubChem CID: ', cid2,
                            '<br> Superclass: ', superclass2,
                            '<br> Class: ', class,
                            '<br> Subclass: ', subclass)) %>%
  layout(scene = list(xaxis = list(title = 'tSNE1'),
                      yaxis = list(title = 'tSNE2'),
                      zaxis = list(title = 'tSNE3')),
         legend = list(itemsizing = "constant"))
p
```



### t-SNE plot of the bitter (in yellow) chemical space, hop-derived compounds 1-11 are coloured in black

```{r, fig.width=12, fig.height=12}
p <- plot_ly(daten2, x = ~X1, y = ~X2, z = ~X3, color = ~group.x,
             colors = c("#ECE471", "black")) %>%
  add_markers(marker = list(size = 5, opacity = 1), hoverinfo = "text",
              text = ~paste("Compound Name: ", str_wrap(IUPACName, width = 40),
                            '<br> PubChem CID: ', cid2,
                            '<br> Hop compound nr.: ', label)) %>%
  layout(scene = list(xaxis = list(title = 'tSNE1'),
                      yaxis = list(title = 'tSNE2'),
                      zaxis = list(title = 'tSNE3')),
         legend = list(itemsizing = "constant"))
p
```

# Chemical space of FooDB, BitterDB and DrugBank


```{r}
data <- read.delim(here("data", "export_tsne.txt"), sep = "\t", header = TRUE, stringsAsFactors=FALSE, quote="")
```



### t-SNE plot of FooDB, BitterDB and DrugBank, coloured according to database occurrence

```{r, warning=FALSE, fig.width=12, fig.height=12}
p <- plot_ly(data, x = ~t.SNE.X.4, y = ~t.SNE.Y.4, z = ~t.SNE.Z.4, color = ~group, colors = "Set1") %>%
  add_markers(marker = list(size = 2, opacity = 0.5), hoverinfo = "text",
              text = ~paste("Compound Name: ", str_wrap(name, width = 40),
                            '<br> Database: ', group)) %>%
  layout(scene = list(xaxis = list(title = 'tSNE1'),
                     yaxis = list(title = 'tSNE2'),
                     zaxis = list(title = 'tSNE3')),
         legend = list(itemsizing = "constant"))
p
```


### t-SNE plot of FooDB, BitterDB and DrugBank, coloured according to ClassyFire chemical superclasses

```{r, fig.width=12, fig.height=12, warning=FALSE}
p <- plot_ly(data, x = ~t.SNE.X.4, y = ~t.SNE.Y.4, z = ~t.SNE.Z.4, color = ~superclass) %>%
  add_markers(marker = list(size = 2, opacity = 0.5), hoverinfo = "text",
              text = ~paste("Compound Name: ", str_wrap(name, width = 40),
                            '<br> Superclass: ', superclass,
                            '<br> Class: ', class,
                            '<br> Subclass: ', subclass)) %>%
  layout(scene = list(xaxis = list(title = 'tSNE1'),
                     yaxis = list(title = 'tSNE2'),
                     zaxis = list(title = 'tSNE3')),
         legend = list(itemsizing = "constant"))
p
```



### t-SNE plot of FooDB, BitterDB and DrugBank highlighting hop-derived compounds 1-11

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=12}
hop <- daten2 %>%
       filter(group.x == "hop-derived bitter compounds") %>%
       select(InChIKey, label, group.x)

data2 <- left_join(data, hop, by = "InChIKey")
data2$group.x[is.na(data2$group.x)] <- "bitter compounds"

p <- plot_ly(data2, x = ~t.SNE.X.4, y = ~t.SNE.Y.4, z = ~t.SNE.Z.4, color = ~group.x,
             colors = c("#ECE471", "black")) %>%
  add_markers(marker = list(size = 2, opacity = 0.5), hoverinfo = "text",
              text = ~paste("Compound Name: ", str_wrap(name, width = 40),
                            '<br> Hop compound nr.: ', label)) %>%
  layout(scene = list(xaxis = list(title = 'tSNE1'),
                     yaxis = list(title = 'tSNE2'),
                     zaxis = list(title = 'tSNE3')),
         legend = list(itemsizing = "constant"))
p
```


